---
title: "Demand Data"
author: "Kat Kyuchukova"
date: "3/23/2019"
output: html_document
---

```{r}
library(tidycensus)
library(tidyverse)
```

Setting up API key
```{r}
## define census API key and set it with census_api_key function

api_key <- "8fa9f8cc6afff6f16cfa41d92974f406a35636a6"
census_api_key(api_key)

## check API key

Sys.getenv("CENSUS_API_KEY")
```

```{r}
v15 <- load_variables(2016, "acs5", cache = TRUE)
v15_race <- v15 %>%
  filter(str_detect(label, fixed("race", 
                                ignore_case = TRUE)))
v15_income <- v15 %>%
  filter(str_detect(label, fixed("income", 
                                ignore_case = TRUE)))

v15_children<- v15 %>%
  filter(str_detect(label, fixed("number of children", 
                                ignore_case = TRUE)))


```

Variable names:
Median income, total: B06011_001
B23020_001: MEAN USUAL HOURS WORKED IN THE PAST 12 MONTHS FOR WORKERS 16 TO 64 YEARS

B08302_001:B08302_015 -- Time leaving home to go to work
B08602_001:015 -- Time arriving at work from home
C24070_001:084 -- INDUSTRY BY CLASS OF WORKER FOR THE CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
B23008_001:B23008_027 -- AGE OF OWN CHILDREN UNDER 18 YEARS IN FAMILIES AND SUBFAMILIES BY LIVING ARRANGEMENTS BY EMPLOYMENT STATUS OF PARENTS
B09002_001:B09002_020 -- OWN CHILDREN UNDER 18 YEARS BY FAMILY TYPE AND AGE
B08303_001:B08303_013 -- travel time to work
C02003_002:C02003_019-- detailed race, Estimate!!Total!!Population of one race!!White/etc.

```{r}
test <- get_acs(geography = "tract",
                     variables = "B08303_001", 
                     survey = "acs5", 
                     state= "MA"
                     )
```






