---
title: "Demand Data"
author: "Kat Kyuchukova"
date: "3/23/2019"
output: html_document
---

```{r}
library(tidycensus)
library(tidyverse)
library(tigris)
library(sf)
library(mosaic)
library(mapview)
library(leaflet)
library(leaflet.extras)
```

Setting up API key
```{r}
## define census API key and set it with census_api_key function

api_key <- "8fa9f8cc6afff6f16cfa41d92974f406a35636a6"
census_api_key(api_key)

## check API key

Sys.getenv("CENSUS_API_KEY")
```

```{r}
v15 <- load_variables(2016, "acs5", cache = TRUE)

v15_race <- v15 %>%
  filter(str_detect(label, fixed("race", 
                                ignore_case = TRUE)))
v15_income <- v15 %>%
  filter(str_detect(label, fixed("income", 
                                ignore_case = TRUE)))

v15_children<- v15 %>%
  filter(str_detect(label, fixed("number of children", 
                                ignore_case = TRUE)))


```

Variable names:
B19013_001: median household income
B23020_001: MEAN USUAL HOURS WORKED IN THE PAST 12 MONTHS FOR WORKERS 16 TO 64 YEARS

B08302_001:B08302_015 -- Time leaving home to go to work
B08602_001:015 -- Time arriving at work from home
C24070_001:084 -- INDUSTRY BY CLASS OF WORKER FOR THE CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER
B23008_001:B23008_027 -- AGE OF OWN CHILDREN UNDER 18 YEARS IN FAMILIES AND SUBFAMILIES BY LIVING ARRANGEMENTS BY EMPLOYMENT STATUS OF PARENTS
B09002_001:B09002_020 -- OWN CHILDREN UNDER 18 YEARS BY FAMILY TYPE AND AGE
B08303_001:B08303_013 -- travel time to work
C02003_002:C02003_019-- detailed race, Estimate!!Total!!Population of one race!!White/etc.

##Example: Time Leaving home
```{r}
leavinghome_vars <- c("B08302_001", "B08302_002", "B08302_003", "B08302_004", "B08302_005", "B08302_006", "B08302_007", "B08302_008", "B08302_009", "B08302_010", "B08302_011", "B08302_012", "B08302_013", "B08302_014", "B08302_015")

leavinghome_before7am_vars <- c("B08302_002", "B08302_003", "B08302_004", "B08302_005", "B08302_006")
leavinghome_am_vars <- c("B08302_002", "B08302_003", "B08302_004", "B08302_005", "B08302_006")

leavinghome_before7 <- get_acs(geography = "tract",
                     variables = leavinghome_before7am_vars,
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)


 leavinghome_after4 <- get_acs(geography = "tract",
                     variables = "B08302_015",
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

plot(leavinghome_before7["estimate"])

ggplot(leavinghome_before7, aes(fill=estimate)) + geom_sf()
ggplot(leavinghome_after4, aes(fill=estimate)) + geom_sf()

m <- mapview(leavinghome_before7, zcol="estimate")
m@map

m2 <- mapview(leavinghome_after4, zcol="estimate")
m2@map

```



##Median income
```{r}
med_income <- get_acs(geography = "tract",
                     variables = "B19013_001",
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

mapviewGetOption("basemaps")
mapviewOptions(basemaps = c("CartoDB.Positron","OpenStreetMap"),
               raster.palette = grey.colors,
               vector.palette = colorRampPalette(c("snow", "cornflowerblue", "grey10")),
               layers.control.pos = "topright")

income_map <- mapview(med_income, zcol="estimate")
income_map@map

  

```

Travel time to work

```{r}
#B08303_001:B08303_013

travel_time_vars <- c("B08303_001", "B08303_002", "B08303_003", "B08303_004", "B08303_005", "B08303_006", "B08303_007", "B08303_008", "B08303_009", "B08303_010", "B08303_011", "B08303_012", "B08303_013")

travel_time <- get_acs(geography = "tract",
                     variables = travel_time_vars,
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

travel_time_long <- travel_time %>%
  spread(variable, estimate)

travel_time_long[is.na(travel_time_long)] <- 0

travel_time_long <- travel_time_long %>%
  group_by(GEOID, NAME) %>%
  summarise(under_30 = sum(B08303_001 + B08303_002 + B08303_003 + B08303_004 + B08303_005 + B08303_006 + B08303_007),
            thirty_to_sixty = sum(B08303_008 +B08303_009 + B08303_010 + B08303_011),
            sixty_to_ninety = sum(B08303_012),
            over_ninety = sum(B08303_013)) %>%
  gather(key = "travel_mins", value= "number_of_people", -GEOID, -NAME, -geometry) %>%
  arrange(NAME)

mapviewGetOption("basemaps")
mapviewOptions(basemaps = c("CartoDB.Positron","OpenStreetMap"),
               raster.palette = grey.colors,
               vector.palette = colorRampPalette(c("snow", "cornflowerblue", "grey10")),
               layers.control.pos = "topright")

travel_time_map <- mapview(travel_time_long, zcol="travel_mins")
travel_time_map@map
```


###MAPPING PARENTAL DEMAND
```{r}
#Map children under 11 by census tract in Boston who are at least 200 FPL (2.00 or higher)
childrenunder11vars <- c("B17024_011", "B17024_012", "B17024_013", "B17024_014", "B17024_024", "B17024_025", "B17024_026", "B17024_027")

children_under_11 <- get_acs(geography = "tract",
                     variables = childrenunder11vars,
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

children_under_11_grouped <- children_under_11 %>%
  group_by(GEOID, NAME) %>%
  summarise(num_children = sum(estimate))

#Multiply this by the percentage of children under 6 with all parents in the labor force, 1 for single-parent household, 2 for two-parent household in each census tract in Boston.

laborforcevars <- c("B23008_002","B23008_004", "B23008_010", "B23008_013")

laborforce_under6  <- get_acs(geography = "tract",
                     variables = laborforcevars,
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

laborforce_under6 <- laborforce_under6 %>%
  group_by(GEOID, NAME) %>%
  summarise(laborforcechildren = sum(estimate))

under6total <- get_acs(geography = "tract",
                     variables = "B23008_002",
                     survey="acs5",
                     state= "MA",
                     county="Suffolk",
                geometry=TRUE)

childrenunder6 <- st_join(laborforce_under6, under6total, by="NAME")

childrenunder6 <- inner_join(laborforce_under6 %>% as.data.frame(), under6total %>% as.data.frame(), by = "NAME")
childrenunder6 <- childrenunder6 %>% 
  st_sf(sf_column_name = 'geometry.x')

childrenunder6 <- childrenunder6 %>%
  mutate(percent_children_laborforce = laborforcechildren/estimate)

#Joining children data with other datasets


```



